---
output: github_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gaborioRtools

<!-- badges: start -->
<!-- badges: end -->

This is a personal package with useful functions for data wrangling, analysis, and teaching R programming, specifically designed for quantitative methods courses in International Relations at Javeriana University.

## Features

- **Data Management Functions**: Tools for cleaning, recoding, and restructuring data
- **Analysis Functions**: Statistical utilities like summary_se for enhanced statistical summaries
- **Teaching Materials**: R Markdown documents for classes and functions to easily open them

## Installation

You can install the development version of gaborioRtools from [GitHub](https://github.com/) with:

```{r eval=FALSE}
# Install devtools if you haven't already
install.packages("devtools")

# Install the package
devtools::install_github("gaborio/gaborioRtools")
```

Alternatively, you can use the provided installation script:

```{r eval=FALSE}
# Download and run the installation script
source("https://raw.githubusercontent.com/gaborio/gaborioRtools/main/install_gaborioRtools.R")
```

## Using the Class Materials

This package includes R Markdown documents for teaching R programming. You can open them with:

```{r eval=FALSE}
library(gaborioRtools)

# Open Class 5 materials (Introduction to R - Part 1)
open_class_doc(5)

# Open Class 6 materials (Introduction to R - Part 2)
open_class_doc(6)
```

## Examples

Here are some examples of the data management functions:

```{r eval=FALSE}
library(gaborioRtools)

# Recode values to 0-100 scale
cols_values <- list(
  EMO_11 = c("5", "6. Positiva"),
  PAF_11 = c("5", "6. Honestos")
)
for (col in names(cols_values)) {
  data <- recode_100(data, col, cols_values[[col]])
}

# Convert percentages to proportions
proportions <- divide_by_x(percentages, 100)

# Scale values to 0-100 range
scores_100 <- from_0_to_100(raw_scores)

# Filter columns that have at least one non-NA value
clean_data <- data %>% select(where(not_all_na))
```

## Dependencies

This package requires several R packages, including:

- tidyverse, haven, readxl, knitr
- jtools, interactions, stargazer
- labelled, showtext, summarytools
- And more (see DESCRIPTION file)

These dependencies will be installed automatically when you install the package.
